---
title: "p8105_hw6_zz3308"
author: "Zitao Zhang"
date: "2024-11-24"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(broom)
library(modelr)
library(purrr)
library(tidyr)
```

# Problem 1

```{r, message=FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r}
bootstrap_results <- weather_df %>%
  modelr::bootstrap(n = 5000) %>%
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin, data = df)),
    results = map(models, \(model) {
      r_squared <- glance(model) %>% pull(r.squared)
      coefficients <- tidy(model) %>% pull(estimate)
      log_beta_product <- log(coefficients[1] * coefficients[2])
      tibble(r_squared = r_squared, log_beta_product = log_beta_product)
    })
  ) %>%
  unnest(results)

ci_r_squared <- quantile(bootstrap_results$r_squared, c(0.025, 0.975))
ci_log_beta_product <- quantile(bootstrap_results$log_beta_product, c(0.025, 0.975))

cat("95% CI for r^2:", ci_r_squared, "\n")
cat("95% CI for log(β0 * β1):", ci_log_beta_product, "\n")

ggplot(bootstrap_results, aes(x = r_squared)) +
  geom_histogram(bins = 30, fill = "blue", alpha = 0.7) +
  geom_vline(xintercept = ci_r_squared, color = "red", linetype = "dashed") +
  labs(
    title = "Bootstrap Distribution of r^2",
    x = "r^2",
    y = "Frequency"
  )

ggplot(bootstrap_results, aes(x = log_beta_product)) +
  geom_histogram(bins = 30, fill = "green", alpha = 0.7) +
  geom_vline(xintercept = ci_log_beta_product, color = "red", linetype = "dashed") +
  labs(
    title = "Bootstrap Distribution of log(β0 * β1)",
    x = "log(β0 * β1)",
    y = "Frequency"
  )
```

The blue histogram shows the bootstrap distribution of $r^2$, centered around 0.91, with a 95% confidence interval from 0.89 to 0.93, indicating strong model predictive power.

The green histogram shows the bootstrap distribution of $log(\beta_0 \cdot \beta_1)$, centered around 2.00, with a 95% confidence interval from 1.97 to 2.06, indicating consistency in the coefficient relationship.


# Problem 2

```{r}
url <- "https://github.com/washingtonpost/data-homicides/blob/master/homicide-data.csv?raw=true"
homicide_data <- read.csv(url) %>% janitor::clean_names()
```

```{r}
homicide_data <- homicide_data %>%
  mutate(
    city_state = paste(city, state, sep = ", "),
    victim_age = ifelse(victim_age == "Unknown", NA, victim_age),
    victim_age = as.numeric(victim_age),
    resolved = disposition == "Closed by arrest"
  ) %>%
  filter(
    !(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")), 
    victim_race %in% c("White", "Black"),
    !is.na(victim_age),
    !is.na(victim_race),
    !is.na(victim_sex)
  )
```

```{r}
baltimore_data <- homicide_data %>% filter(city_state == "Baltimore, MD")

baltimore_model <- glm(
  resolved ~ victim_age + victim_sex + victim_race,
  data = baltimore_data,
  family = binomial
)

baltimore_results <- tidy(baltimore_model, exponentiate = TRUE, conf.int = TRUE) %>%
  filter(term == "victim_sexMale")

print(baltimore_results)
```

```{r, warning=FALSE}
city_results <- homicide_data %>%
  group_by(city_state) %>%
  nest() %>%
  mutate(
    models = map(data, ~ glm(
      resolved ~ victim_age + victim_sex + victim_race,
      data = .x,
      family = binomial
    )),
    results = map(models, ~ tidy(.x, exponentiate = TRUE, conf.int = TRUE))
  ) %>%
  unnest(results) %>%
  filter(term == "victim_sexMale") %>%
  select(city_state, estimate, conf.low, conf.high)

print(city_results)
```

```{r}
ggplot(city_results, aes(x = reorder(city_state, estimate), y = estimate)) +
  geom_point(color = "blue") +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, color = "red") +
  coord_flip() +
  labs(
    title = "Adjusted Odds Ratios (Male vs Female Victims) for Resolved Homicides",
    x = "City",
    y = "Adjusted Odds Ratio (Male vs Female)"
  ) +
  theme(axis.text.y = element_text(size = 6))
```

The plot illustrates the adjusted odds ratios (ORs) for resolving homicides involving male versus female victims across cities, with confidence intervals highlighting variability in the estimates. While some cities show statistically significant differences (confidence intervals entirely above or below 1), many have wide intervals reflecting uncertainty, likely due to limited data. Cities with narrower intervals, such as Baltimore, provide more reliable estimates, whereas cities like Albuquerque indicate a need for further data review. Overall, there is substantial heterogeneity across cities in the resolution likelihood based on victim sex.


# Problem 3

```{r}
birthweight_data <- read.csv("data/birthweight.csv") %>% janitor::clean_names()

birthweight_data <- birthweight_data %>%
  mutate(
    babysex = factor(babysex, levels = c(1, 2), labels = c("Male", "Female")),
    frace = factor(frace, 
                   levels = c(1, 2, 3, 4, 8, 9),
                   labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")),
    mrace = factor(mrace, 
                   levels = c(1, 2, 3, 4, 8),
                   labels = c("White", "Black", "Asian", "Puerto Rican", "Other")),
    malform = factor(malform,
                     levels = c(0, 1),
                     labels = c("absent", "present"))) %>%
  drop_na()
```

```{r}
proposed_model <- lm(
  bwt ~ blength + bhead + gaweeks + wtgain + ppbmi + delwt + malform,
  data = birthweight_data
)

birthweight_data <- birthweight_data %>%
  add_predictions(proposed_model) %>%
  add_residuals(proposed_model)

ggplot(birthweight_data, aes(x = pred, y = resid)) +
  geom_point(alpha = 0.6) +
  geom_smooth(se = FALSE, color = "blue") +
  labs(
    title = "Residuals vs Fitted Values",
    x = "Fitted Values",
    y = "Residuals"
  ) +
  theme_minimal()
```

I choose Hypothesis-Driven Selection, I started by identifying key variables likely to influence birthweight (`bwt`) based on domain knowledge and biological plausibility. These include:

Baby Characteristics: `blength` (baby’s length) and `bhead` (head circumference) are direct indicators of fetal size./
Pregnancy-Related Factors: `gaweeks` (gestational age) strongly correlates with birthweight, as longer gestation generally allows more growth./
Maternal Health: `wtgain` (weight gain during pregnancy) and `delwt` (weight at delivery) reflect maternal health and nutrition, including `malform`, which are critical for fetal growth./
Maternal Pre-pregnancy Condition: `ppbmi` (pre-pregnancy BMI) serves as a baseline indicator of the mother's health and nutritional status.

The residual plot shows residuals scattered around zero, indicating no major violations of linearity. However, there is a slight curve in the residuals, particularly at lower fitted values, suggesting some potential non-linearity in the model. Additionally, the spread of residuals appears relatively consistent, with no strong evidence of heteroscedasticity.

```{r}
model1 <- lm(bwt ~ blength + gaweeks, data = birthweight_data)

model2 <- lm(bwt ~ bhead * blength * babysex, data = birthweight_data)
```

```{r}
set.seed(123)

cv_df <- birthweight_data %>% 
  crossv_mc(100) %>%
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  ) %>%
  mutate(
    proposed_model = map(train, \(df) lm(bwt ~ blength + bhead + gaweeks + wtgain + ppbmi + delwt + malform, data = df)),
    model1 = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)),
    model2 = map(train, \(df) lm(bwt ~ bhead * blength * babysex, data = df))
  ) %>%
  mutate(
    rmse_proposed = map2_dbl(proposed_model, test, \(mod, df) rmse(model = mod, data = df)),
    rmse1 = map2_dbl(model1, test, \(mod, df) rmse(model = mod, data = df)),
    rmse2 = map2_dbl(model2, test, \(mod, df) rmse(model = mod, data = df)))

cv_df %>% 
  select(starts_with("rmse")) %>%
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>%
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin() +
  labs(
    title = "rmse comparison of models",
    x = "model",
    y = "rmse"
  )
```

The violin plot shows that the Proposed Model performs best, with the lowest median RMSE and the most consistent results across cross-validation splits. Model 2 (`bhead` * `blength` * `babysex`) performs better on average than Model 1 (`blength` + `gaweeks`), as it has a lower median RMSE, but its wider spread indicates higher variability. Model 1 is more stable but slightly less accurate on average. Overall, the Proposed Model outperforms both alternatives in accuracy and reliability.

















